Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-05-17 manuscript/revision2/revision2.tex
chores: ensure revision2 images render from images folder
revision: Add SI Appendix folder/files for revision2

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (30).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (31, 32). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and survival (33). The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
(34, 35), while the Gulf subpopulation has individuals from the
well-studied lowland ecotype and the phenotypically intermediate coastal
ecotype (33).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included
(@SI-info) (36). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage) developed by (36), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, this method circumvents statistical biases in
detecting genetic effects with the same or opposite signs (37). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which can flexibly capture many modes of covariance across gardens
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
qualitative categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfall amounts are uncorrelated,
covariance will be near zero; if there is rank changing (e.g.~high
rainfall before flowering of A at MO but not MI, and high rainfall
before flowering for B at MI but not MO), this covariance will be near
negative one. We specified multiple covariance structures for the
canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the mash model.

@@

We used a greedy algorithm to select covariance matrices from each
model's set that significantly improved the model likelihood. Five
GxWeather covariance structures were selected, zero to two per
subpopulation: two for the start of vegetative growth, and three for
flowering date. Of these five, three had mass on them in mash models of
the strong effects (Figure~\ref{fig-covar} A-B). Two of these three
matrices had negative covariances between sites between Texas and North
gardens (Figure~\ref{fig-covar} A), while one had all positive or
near-zero covariances. SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash
models of the timing of vegetative growth in the Midwest subpopulation
and Both subpopulations did not include this GxWeather covariance; the
Midwest had no weight on any GxWeather covariance, while Both
subpopulations had non-zero weights on two additional GxWeather
covariance types, average temperature one day prior to green-up, and the
day length change in seconds in the day prior to green-up
(Figure~\ref{fig-covar} B). The average temperature covariance matrix
had negative covariances between Texas and North gardens, though not as
strong as the negative phenotypic correlations seen in Both
subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
the Gulf subpopulation and Both subpopulations had mass on any GxWeather
covariance matrices (Figure~\ref{fig-covar} C) for vegetative growth.

@@

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} A; (\textbf{SI-fig-gardens?})). 78.7\% of
pairwise comparisons between North and Texas gardens had a difference in
sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
comparisons had a difference in sign, respectively
(Figure~\ref{fig-effects}). The majority of pairwise effects for greenup
for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
subpopulations were the same sign, and effects most often differed in
magnitude between gardens within and between the regions
(Figure~\ref{fig-effects} X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects}). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X),
while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign
(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
the southernmost garden, and all other gardens
((\textbf{SI-fig-gardens?})). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

@@

\begin{table}[t!]
\centering
\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices tested \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

\end{table}

@@

\includegraphics{Figure_2_recolor_covariance_matrices_and_posterior_weights.png}

@@

\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
specified in mash and the posterior weights placed on all covariance
matrices. (a) Left column: Five example hypothesized covariance matrices
specified for the green-up date or flowering date phenotype; these
matrices were created from environment-specific correlations across
eight common gardens, and are described in Table XXX. Common gardens are
arranged in latitudinal order within the matrices. Right column: Five
example canonical covariance matrices. Canonical matrices (purple) have
simple interpretations, such as equal effects across all common gardens,
or effects specific to a single common garden. (b,d) Total posterior
weight placed on each covariance matrix type specified for (b) green-up
date and (d) flowering date mash models, within and between two genetic
subpopulations. Hypothesized covariance matrices (green). Covariance
matrices included in mash that had zero posterior weight in all three
mash runs on the genetic subpopulations, such as the identity matrix,
are not shown. (c,e) Total posterior weight placed on covariance
matrices that were hypothesized or canonical, for the (c) green-up date
phenotype and (e) flowering date phenotype.}

@@

\includegraphics{Figure_3_GxE.png}

@@

\includegraphics{Figure_4_QTL_Overlaps.png}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (32, 34, 37). We scored the
start of vegetative growth as the day of the year when 50\% of the
tiller area of the crown of the plant cut the previous year had green
growth. The start of reproductive growth, or flowering date, was the day
of the year when 50\% of the plant tillers had panicles undergoing
anthesis. We scored green-up and flowering as day of the year, then
linked these dates to multiple weather-based environmental factors
measured daily at each common garden (SI Appendix, Section S1, Table
S1).

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (38), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both'). Analyses to determine narrow-sense
heritability (h\textsuperscript{2}) for green-up and flowering were done
using linear mixed models and followed (33). Details on these models can
be found in (SI Appendix, Section S3,S4).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for the two traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from
correlations in environmental cues in the green-up or flowering date
windows for the three populations (SI Appendix, Table S1, Section S1).
These covariance matrices represent correlations between identical
genotypes drawn from a specific population at pairs of common gardens;
covariances near one mean that the population has a strong, positive
linear relationship in individual responses at that pair of gardens,
while covariances near zero mean that there is no relationship within
the population for individual responses at that pair of gardens. Mash
SNP effects will undergo strong shrinkage towards one another in the
first case, and little shrinkage in the second case. Mash also generates
data-driven covariance matrices corresponding to major patterns of SNP
effects present in the data. We generated six data-driven matrices per
mash run, five produced by singular value decomposition (SVD) of an
overall matrix.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy
(e.g.~(37)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (26, 27), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (37). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

@@

>>>>

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (30).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (31, 32). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and overwinter survival (33). The Midwest genetic subpopulation is
primarily composed of individuals from the well-studied upland
switchgrass ecotype (34, 35), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (33).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(Table 1) and additional ways phenological timings might vary by site
(SI Appendix, Table S1, Section SX), then jointly re-estimate genetic
effects on these timings at all eight sites using the set of these
covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section SX) (36). We use the
Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (36), to refine effect size estimates from genome-wide
association (GWAS) conducted on individuals at each of the eight sites.
\emph{mash} allows us to identify and specify multiple covariance
structures among genetic effect estimates across sites, including
structures that represent covariance in weather variables of interest.
Importantly, this method circumvents statistical biases in detecting
genetic effects with the same or opposite signs (37). To confirm our
genetic mapping of GxWeather, we compare our posterior genetic effect
estimates from \emph{mash} to mapping results from an outbred pseudo-F2
cross grown at the same sites. Our analyses allow us to describe the
weather cues and genetic variation affecting phenology in two divergent
natural populations of switchgrass.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which can flexibly capture many modes of covariance across gardens
(Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We specified three
qualitative categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1; SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfall amounts are uncorrelated,
covariance will be near zero; if there is rank changing (e.g.~high
rainfall before flowering of A at MO but not MI, and high rainfall
before flowering for B at MI but not MO), this covariance will be near
negative one. We specified multiple covariance structures for the
canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the mash model.

@@

We used a greedy algorithm to select covariance matrices from each
model's set that significantly improved the model likelihood. Five
GxWeather covariance structures were selected, zero to two per
subpopulation: two for the start of vegetative growth, and three for
flowering date. Of these five, three had mass on them in mash models of
the strong effects (Figure~\ref{fig-covar} A-B). Two of these three
matrices had negative covariances between sites between Texas and North
gardens (Figure~\ref{fig-covar} A), while one had all positive or
near-zero covariances. SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash
models of the timing of vegetative growth in the Midwest subpopulation
and Both subpopulations did not include this GxWeather covariance; the
Midwest had no weight on any GxWeather covariance, while Both
subpopulations had non-zero weights on two additional GxWeather
covariance types, average temperature one day prior to green-up, and the
day length change in seconds in the day prior to green-up
(Figure~\ref{fig-covar} B). The average temperature covariance matrix
had negative covariances between Texas and North gardens, though not as
strong as the negative phenotypic correlations seen in Both
subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
the Gulf subpopulation and Both subpopulations had mass on any GxWeather
covariance matrices (Figure~\ref{fig-covar} C) for vegetative growth.

@@

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} A; SI Appendix, Fig SX). 78.7\% of pairwise
comparisons between North and Texas gardens had a difference in sign,
while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
had a difference in sign, respectively (Figure~\ref{fig-effects}). The
majority of pairwise effects for greenup for the Midwest
(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
same sign, and effects most often differed in magnitude between gardens
within and between the regions (Figure~\ref{fig-effects} X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects}). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X),
while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign
(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
the southernmost garden, and all other gardens (SI Appendix, Fig SX).
Similarly, more effects that differed in magnitude included gardens in
the Texas region (52.3-55.9\%), and most effect pairs in the North
region were not distinguishable (91.5\%).

@@

\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

\end{table}
\clearpage
\twocolumn

@@

\includegraphics{images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png}

@@

\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
specified in mash and the posterior weights placed on all covariance
matrices. (a) Left column: Five example hypothesized covariance matrices
specified for the green-up date or flowering date phenotype; these
matrices were created from environment-specific correlations across
eight common gardens, and are described in Table 1. Common gardens are
arranged in latitudinal order within the matrices. Right column: Five
example canonical covariance matrices. Canonical matrices (purple) have
simple interpretations, such as equal effects across all common gardens,
or effects specific to a single common garden. (b,d) Total posterior
weight placed on each covariance matrix type specified for (b) green-up
date and (d) flowering date mash models, within and between two genetic
subpopulations. Hypothesized covariance matrices (green). Covariance
matrices included in mash that had zero posterior weight in all three
mash runs on the genetic subpopulations, such as the identity matrix,
are not shown. (c,e) Total posterior weight placed on covariance
matrices that were hypothesized or canonical, for the (c) green-up date
phenotype and (e) flowering date phenotype.}

@@

\includegraphics{images/Figure_3_GxE.png}

@@

\includegraphics{images/Figure_4_QTL_Overlaps.png}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (35, 39, 40). We scored the
start of vegetative growth as the day of the year when 50\% of the
tiller area of the crown of the plant cut the previous year had green
growth. The start of reproductive growth, or flowering date, was the day
of the year when 50\% of the plant tillers had panicles undergoing
anthesis. We scored green-up and flowering as day of the year, then
linked these dates to multiple weather-based environmental factors
measured daily at each common garden (SI Appendix, Section S1, Table
S1).

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (41), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both'). Analyses to determine narrow-sense
heritability (h\textsuperscript{2}) for green-up and flowering were done
using linear mixed models and followed (33). Details on these models can
be found in (SI Appendix, Section S3,S4).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(36). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for the two traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from
correlations in environmental cues in the green-up or flowering date
windows for the three populations (Table 1; SI Appendix, Table S1,
Section S1). These covariance matrices represent correlations between
identical genotypes drawn from a specific population at pairs of common
gardens; covariances near one mean that the population has a strong,
positive linear relationship in individual responses at that pair of
gardens, while covariances near zero mean that there is no relationship
within the population for individual responses at that pair of gardens.
Mash SNP effects will undergo strong shrinkage towards one another in
the first case, and little shrinkage in the second case. Mash also
generates data-driven covariance matrices corresponding to major
patterns of SNP effects present in the data. We generated six
data-driven matrices per mash run, five produced by singular value
decomposition (SVD) of an overall matrix.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy (e.g.
(39)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (27, 29), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (37). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

@@

\bibitem[\citeproctext]{ref-kortefarlow2013}
\CSLLeftMargin{41. }%
\CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
trait analysis with GWAS: A review. \emph{Plant methods} \textbf{9}, 29
(2013).}
